Appendix B of Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-12-14 manuscript/bef_stability/appendixB.Rmd
Last edits before submission

---
title: Appendix of Species richness and network structure jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   -  \newcommand{\beginsupplement}{\setcounter{table}{2}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

The fish community sampling took place in a variety of stream size, shape and
depth. This diversity constraints the type of protocol that can take place.  For
the small streams (depth < 0.7m across the entire width), the complete protocol
is used. It consists in placing anodes every X meters across the width of the
river. The fishermen made from one to three row
of fishing. We kept only the first row for analysis to homogenize the sampling
effort. For deeper and wider streams, the sampling was partial over the bank and
the channel and performed by boat and by foot (along the bank). We kept only
sampling done over the bank because channel sampling was not always done. Two
types of partial protocols takes place: partial over bank and partial by points.
The former is continuous while the latter is discrete. Furthermore, in partial
by point protocol the channel can be sampled as well. We kept only the partial
by point fish sampling which all the sampling point were done over the bank to
homogenize the sampling. Then, we merged partial over bank and partial over
point protocols. Some sampling stations had mix characteristics in which mix of
partial and complete protocols were applied depending of the depth of the river
at the time of sampling.

Heterogeneity in the sampling period may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in sampling time, only one sampling per year was retained if there
was two in a given year, the one with with the most species and then individual
collected. Finally, most of the fish sampling occurred in September (Median:
mid-August) and the median standard deviation of the sampling month was inferior
to one (0.7) within station. Additionally, we removed sampling events, in a
given station, which the sampled length along the stream was outside the median
length sampled more or less 30%. Finally, we standardized the biomass record
and species richness by the sampled surface to get rid of sampling effort
effects.

@@

During the electric fishing, and before that fish body size are measured, they
are regrouped by batches. Each batches contains one species but there can be
several batches by species. Body size assessment are different according to the
batches. There are four types of batches: N, I, S/L and G. The batches N and
I contains the biggest individuals, there are normally less numerous and then
the body size of the individuals of those batches is recorded. The batches S/L
contains fishes of intermediate sizes. Inside each S/L batches, the operator
tries to make batches as homogeneous as possible in body size. Then, the body
size of a subsample of X individuals are recorded. The G batches contain
individuals of low body size, with care that body size is as homogeneous as
possible in the batches. Then body size of the shortest and the longest
individuals are recorded.

@@

```{r}
body_size_caption <- paste0(
  "Body size inference of the fishes. ",
  "(1) Sampled fishes were attributed to batch type according to their size and
  the number of individuals. A batch contains one species only. (2) In the batch
  type N and I, the body size of all the individuals was recorded. In the
  batches, S/L, the body size of a subsample of the individuals was recorded.
  Finally, only the body size of the littlest and of the tallest individual was
  recorded in the G batches. ",
  "Then, (3) we inferred the body size distribution in the S/L and G batches by
  assuming that it followed a normal distribution. The mean and the standard
  deviation were estimated with the measured subsample."
) 
```

```{r p-body-size, fig.cap=body_size_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "length_inference.pdf"))
```

With the information contained in the batch record, we inferred the body size of
the fishes in the community. No inference were needed for the N and I batches
because the body size of the individuals were recorded. The body size ($l$) of the
fishes in the S/L and G were inferred under the assumption that its distribution
followed a normal law ($l ~ N(\mu , \sigma)$).
For a S/L batch, the mean $\mu$ and the standard deviation $\sigma$ were simply
the average and the standard deviation of the body size of the subsample
recorded. For a G batch, the average was approximated by 
$\hat{\mu} = (l_{min} + l_{max}) 2^{-1}$ and the standard deviation by
$\hat{\sigma} = (l_{min} + l_{max}) 4^{-1}$. Together with the estimations of
the mean and the standard deviation, the number of individuals in a batch and
assumption of a normal distribution, we simulated the body size distribution in
S/L and G batches. To avoid simulation of unrealistic body size, we used a
truncated normal distribution bounded at [0.05; .95] for the density of
probability of the S/L batches, and bounded at [$l_{min}$; $l_{max}$] for the G
batches. Body size distribution were generated using the R function
`truncdist::rtrunc`.

We removed batches that did not satisfied the following conditions. First, we
removed batches which the total number of individuals was not recorded. In lot
"G", we removed batches were  $l_{min} \geq l_{max}$ or if at least one of the
measurement was missing. Additionally, G batches were removed when number of
individuals fell below five, as G batches normally contained more individuals.
In the S/L lot, we removed batches were the number of individuals in the
recorded individuals fell below 10. In sum, those sanity check allowed to
exclude potentially problematic batches. Finally, in the dataset used in the
analysis, `r na_length` individuals had missing length over 
`r nrow_fish_length` individuals.

@@

We removed rare fish species, species from lake, crayfish species, and migratory
species. In total, we removed `r nrow(removed_ind)` over 
`r nrow(length_sem_analysis)` individuals and `r length(removed_sp)`.

During fish sampling, some species are identified based on morphotype but they
belong to the same species based on fecundation criteria.  We merged those
morphotype species into one species. There was `r
length(unique(names(sp_to_replace())))` to merge which belonged to `r
length(unique(sp_to_replace()))` species.

@@

Diet information were obtained from fishbase.org and literature (Appendix A, Table S2). The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had as strict piscivorous
life stage. Fish species were largely omnivorous.

@@

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
resource nodes were determined by literature. For example, phytobenthos fed on
detritus, while zooplankton fed on zooplankton and phytoplankton. Fish-fish
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies [@brose_predator_2019; @poisot_structure_2016;
@schneider_body_2012].  For a given trophic species, we first determined if its
size class end up in a life stage containing piscivorous diet, i.e. if there was
an overlap between them. We were more liberal for piscivory determination to not
miss potential fish-fish trophic interactions which were the main interactions
in our community. If the trophic species was piscivorous, we determined its prey
size range from 3% to 45% of the median of its size range. A trophic interaction
was set for each trophic species whose median of its size class fell into the
prey range of the predator trophic species. To sum up, the metaweb was obtained
with allometric feeding interactions (1) between trophic species, (2) between
trophic species and resource nodes and (3) between resource nodes.

@@

The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each
recording station, we applied a moving window averaging (more details in
supplementary material) to get rid of seasonal effects. For temperature, the
moving was performed on daily average temperature. Then we took the annual
average of each year and each variable (i.e. one value by year and by station).

Several characteristics of the streams make the interpolation more complicated
than along terrestrial surface. Interpolations in terrestrial areas are roughly
based on euclidean distance. Euclidean distance is more complicated to consider
in streams because it had the follow the path of the streams. Furthermore, two
streams can be close by but not connected and the value of their variables could
be not related or a only a few. For those reason, we used stream spatial network
methods to interpolate temperature, flow and BOD.

We reconstructed the stream network and the connection between sites with the
`openSTARS` R package [@kattwinkel_preparing_2020]. The reconstruction was done with a raster of digital
elevation model at a resolution of 250 meters [(IGN)](https://wxs.ign.fr/jvam1hsjm11u8voorw81v2xb/telechargement/prepackage/BDALTI-75M_PACK_FXX_2018-01-24%24BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15/file/BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15.7z)
and a theoretical hydrologic network [@pella_reseau_2012].

The procedure was done by hydrographic basin because of convergence problem
when the interpolation was done at the France scale. We had to make a trade-off
between the number of data (to keep enough high to perform the interpolation)
and the level of dataset splitting. We then merge some hydrographic basin to end
up with four region of interpolation: West, East, South and North. (XXX ADD 
basins merged)

Then we modeled each variable as dependant only on the covariation between sites
and taking in account stream network. To model the autocovariance between sites,
we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their prediction performance. The interpolations
were done with the `SSN` R package 
[@isaak_applications_2014, @hoef_ssn:_2014].

@@

The performance of the interpolation was evaluated with leave one out
cross-validation (already implemented in the SSN package). The value of the
interpolated variable of each station was predicted after removing it from the
dataset.

@@

The CV of biomass over can cover multiple dynamics, including increases,
stable, decreases and a mix of them. Theoretically, however, the CV computed to
quantify stability is often represented with the assumption that the overall
dynamic of the biomass varies around an equilibrium. In order to approach this
condition, we kept only the stations that have met this criteria. To qualify the
biomass trajectories over time, we used the symbolic aggregated approximation
method (SAX) which is commonly used in time series analysis and implemented in
the `jmotif` R package. This method classifies time series rather than
quantifies trends and so is generic enough to encompass the diversity of time
series shapes. The classification necessitates three steps. First, the time
series are normalized (i.e. Substituting the mean and divide by the standard
deviation). The second step is to reduce the dimensionality of the time series
by Piecewise Aggregation Approximation (PAA) method. Each time series is divided
in $N$ frames of equal size that allows to the reduce the dimensionality of the
time series. For each frame, we simply kept the mean value of the time series.
The last step is to qualify each frame into string. Because values of
normalized time series typically follows a normal distribution ($N(0,1)$), a
breakpoints can be computed by dividing the area under the normal distribution
in equal parts. The algorithm attributes a alphabetic letter to each part from
the lowest to the highest ($a$, $b$, $c$, ...). We set three breakpoints, so
$a$, $b$ and $c$. Finally, the trajectory of each time series is classified as
a combination of $a$, $b$, $c$ letters. $abc$ and $cba$ time series represent
respectively monotonic increasing and decreasing trajectories. For data analysis,
we kept only the $bbb$ which represent time series that showed no temporal
tendencies, with respect to our methodology.

@@

```{r}
prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))

@@

pca_caption <- paste0(
  "Rotated axis of PCA on environmental variables. The five components
  explained,", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  "(A) Axis 1 was positively related to the river size, axis 2 was related to
  average temperature and negatively related to the river slope and altitude.
  (B) Axis 3 was negatively related to the CV of stream size.
  Axis 4 was related to the CV of enrichment.
  Finally, axis 5 was related to the CV of flow and average enrichment.")
```

```{r pca-rotated, fig.cap = pca_caption}
p1 <- my_pca_plot(xaxis = "RC1", yaxis = "RC2", ctb_thld = .4)
#p2 <- my_pca_plot(xaxis = "RC3", yaxis = "RC4", ctb_thld = .4)
#p3 <- my_pca_plot(xaxis = "RC4", yaxis = "RC5", ctb_thld = .4)

@@

The five axis of the rotated PCA explained respectively
`r paste0(prop_var_expl_pca, "%", collapse = ", ")` of the variance.
Please see the main text for the interpretation of the axis.

@@

```{r}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 2)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 2), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c(rep("Log base 10", 2), "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  interpretation = c("linked to stream size",
    "linked to average temperature",
    "Total species richness over the temporal series",
    "Median of annual link density", "Median of annual average trophic level", "", "",
    "Median of annual total biomass")) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "Variable description and transformation"
  ) %>%
kable_styling(latex_options = c("striped"))
sem_variable_table
```

### SEM hypothesis

We linked environmental variables both to community structure and stability
components. Rooted in assembly theory, environment filters species that can
establish in a community and thus affect community structure. (+ add specific
hypothesis about physical and climate). (1) We hypothesized that enrichment
(characterized by DBO) and mean annual temperature will increase species turnover
[@friedpetersen_drivers_2020]. (2) We hypothesized environmental forcing can also
decrease biomass stability directly without going by community structure. First,
instability of habitat and environmental factors may in turn drive biomass
instability [@hansen_climate_2013]. Secondly, average environmental conditions
such as enrichment can increase directly CV by increasing species biomass
fluctuations [@tabi_warming_2019]. Increasing temperature was also hypothesized
to decrease stability through starvation resulting from high metabolic activity
[@tabi_warming_2019].

---
# Species richness to community attributes 
---

To reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. In particular, species richness decreased connectance
following results from @thebault_stability_2010; @winemiller_must_1989. But
theoretically, some ecological networks can show the inverse relationship
[@winemiller_must_1989]. (6) We hypothesized also that species richness will
modulate average weighted trophic level because of food web assembly processes.
A community with more fish species can support higher trophic level. Finally, we
set up causal relationships from community attributes to stability components.

---
#  Community attributes to stability
---
We hypothesized that (7) species richness decreased synchrony between species
[@loreau_species_2008] and increased $\overline{CV_{sp}}$
[@tilman_biodiversity_2006]. (8a) We expected that average trophic level had a
positive or negative relationship with synchrony, bottom-up effects should
result in positive whereas top-down in negative relationships. We expected a (8b)
negative relationship between average trophic level and $\overline{CV_{sp}}$ if
predators stabilised prey fluctuations or if the increasing part of
biomass in the predators decreased $\overline{CV_{sp}}$ because of their higher
body mass (if mean increase faster than variance). (9a) Connectance was expected
to decrease or increase $\overline{CV_{sp}}$ and (9b) synchrony because of top
down effects or bottom up effects.

@@

```{r sensitivity, message = FALSE, fig.dim = c(10, 7), fig.cap = "Sensivity analysis with different dataset"}
library(furrr)

>>>>

---
title: Appendix B of Species richness and network structure jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   -  \newcommand{\beginsupplement}{\setcounter{table}{2}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

The fish community sampling took place in a variety of stream size, shape and
depth. This diversity constraints the type of protocol that can take place.  For
the small streams (depth < 0.7m across the entire width), the complete protocol
is used. It consists in placing anodes across the width of the
stream. The fishermen made from one to three row
of fishing. We kept only the first row for analysis to homogenize the sampling
effort. For deeper and wider streams, the sampling was partial over the bank and
the channel and performed by boat and by foot (along the bank). We kept only
sampling done over the bank because channel sampling was not always done. Two
types of partial protocols takes place: partial over bank and partial by points.
The former is continuous while the latter is discrete. Furthermore, in the "partial
by point" protocol, the channel can be sampled as well. We kept only the "partial
by point" sampling in which all the sampling points were done over the bank to
homogenize the sampling. Then, we merged partial over bank and partial by 
point protocols. Some sampled stations had mixed characteristics in which mix of
partial and complete protocols were applied depending of the depth of the stream 
at the time of sampling. In this case, we kept only the protocol that was the
most used.

Heterogeneity in the sampling period may introduce variability that is not related
to ecological processes such as the presence of juveniles. In order to reduce
heterogeneity in sampling time, only one sampling per year was retained if there
was two in a given year, the one with the most species and then individual
collected. Finally, most of the fish sampling occurred in September (Median:
mid-August) and the median standard deviation of the sampling month was inferior
to one (0.7) within stations. Additionally, we removed sampling events, in a
given station, in which the sampled length along the stream was outside the median
length sampled more or less 30%. Finally, we standardized the biomass record
and species richness by the sampled surface to get rid of sampling effort
effects.

@@

During the electric fishing, and before that fish body size are measured, the
fishes are regrouped by batches. Each batch contains one species but there can
be several batches by species. Body size assessment are different according to
the batches. There are four types of batches: N, I, S/L and G. The batches N and
I contains the biggest individuals which are normally less numerous. Then
the body size of all the individuals in the N and I batches is recorded. The batches S/L
contains fishes of intermediate sizes. Inside each S/L batches, the operator
tries to make batches as homogeneous as possible in terms of body size. Then, the body
size of a subsample of at least 20 individuals is recorded. The G batches
contain individuals of small body size, with care that body size is as
homogeneous as possible in a given batches. Then body size of the shortest and the
longest individuals are recorded.

@@

```{r, cache = FALSE}
body_size_caption <- paste0(
  "Body size inference of the fishes. ",
  "(1) Sampled fishes were attributed to batch type according to their size and
  the number of individuals. A batch contains one species only. (2) In the batch
  type N and I, the body size of all the individuals was recorded. In the
  batches, S/L, the body size of a subsample of the individuals was recorded.
  Finally, only the body size of the smallest and of the largest individual were
  recorded in the G batches. ",
  "Then, (3) we inferred the body size distribution in the S/L and G batches by
  assuming a normal distribution. The mean and the standard
  deviation were estimated with the measured subsample."
) 
```

```{r p-body-size, cache = FALSE, fig.cap=body_size_caption}
knitr::include_graphics(mypath("manuscript", "bef_stability", "figs", "length_inference.pdf"))
```

With the information contained in the batch record, we inferred the body size of
the fishes in the community. No inference were needed for the N and I batches
because the body size of all individuals was recorded. The body size ($l$) of the
fishes in the S/L and G was inferred under the assumption that its distribution
followed a normal law ($l ~ N(\mu , \sigma)$).
For a S/L batch, the mean $\mu$ and the standard deviation $\sigma$ were simply
the average and the standard deviation of the body size of the subsample
recorded. For a G batch, the average was approximated by 
$\hat{\mu} = (l_{min} + l_{max}) 2^{-1}$ and the standard deviation by
$\hat{\sigma} = (l_{min} + l_{max}) 4^{-1}$. Together with the estimations of
the mean and the standard deviation, the number of individuals in a batch and
assumption of a normal distribution, we simulated the body size distribution in
S/L and G batches. To avoid simulation of unrealistic body size, we used a
truncated normal distribution bounded at [0.05; .95] for the density of
probability of the S/L batches, and bounded at [$l_{min}$; $l_{max}$] for the G
batches. Body size distribution were generated using the R function
`truncdist::rtrunc`.

We removed batches that did not satisfied the following conditions. First, we
removed batches in which the total number of individuals was not recorded. In lot
"G", we removed batches where $l_{min} \geq l_{max}$ or if at least one of the
measurement was missing. Additionally, G batches were removed when number of
individuals fell below five, as G batches normally contained more individuals.
In the S/L lot, we removed batches in which the number of individuals in the
recorded individuals fell below 10. In sum, those sanity check allowed to
exclude potentially problematic batches. Finally, in the dataset used in the
analysis, `r na_length` individuals had missing length over 
`r nrow_fish_length` individuals.

@@

We removed rare fish species, species from lake, crayfish species, and migratory
species. In total, we removed `r nrow(removed_ind)` over 
`r nrow(length_sem_analysis)` individuals and `r length(removed_sp)` species.

During fish sampling, some species are identified based on morphotype but they
belong to the same species based on fecundation criteria.  We merged those
morphotype species into one species. There was 
`r length(unique(names(sp_to_replace())))` species to merge which belonged to 
`r length(unique(sp_to_replace()))` species.

@@

Diet information were obtained from fishbase.org and literature (Appendix a, Table S2). The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had a strict piscivorous
life stage. Fish species were then largely omnivorous.

@@

A trophic species belonged to a life stage if the median of its body size range
fell in the range of the life stage. In turn, this trophic species has feeding
interactions with the resource nodes of its life stage. Interactions between
trophic interactions between a predator and a prey were determined by body size
ratio as in many studies [@brose_predator_2019; @poisot_structure_2016;
@schneider_body_2012].  For a given trophic species, we first determined if its
size class ends up in a life stage containing piscivorous diet, i.e. if there was
an overlap between them. We were more liberal for piscivory determination to
avoid 
missing potential fish-fish trophic interactions which were the main interactions
in our community. If the trophic species was piscivorous, we determined its prey
size range from 3% to 45% of the median of its size range. A trophic interaction
was set for each trophic species whose median of its size class fell into the
prey range of the predator trophic species. To sum up, the metaweb was obtained
with allometric feeding interactions (1) between trophic species, (2) between
trophic species and resource nodes and (3) between resource nodes.

@@

The time resolution was heterogeneous according to the variable considered
(hourly, daily and monthly for temperature, flow and BOD resp.). For each
recording station, we applied a moving window averaging to get rid of seasonal
effects. For temperature, the moving was performed on daily average temperature.
Then we took the annual average of each year and each variable (i.e. one value
by year and by station).

Several characteristics of the streams make the interpolation more complicated
than along terrestrial surface. Interpolations in terrestrial areas are roughly
based on euclidean distance. Euclidean distance is more complicated to consider
in streams because it had the follow the path of the streams. Furthermore, two
streams can be close by but not connected and the value of their variables could
be not related or a only a few. For those reason, we used stream spatial network
methods to interpolate water temperature, flow and BOD.

We reconstructed the stream network and the connection between stations with the
`openSTARS` R package [@kattwinkel_preparing_2020]. The reconstruction was done with a raster of digital
elevation model at a resolution of 250 meters [(IGN)](https://wxs.ign.fr/jvam1hsjm11u8voorw81v2xb/telechargement/prepackage/BDALTI-75M_PACK_FXX_2018-01-24%24BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15/file/BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15.7z)
and a theoretical hydrologic network [@pella_reseau_2012].

The procedure was done by hydrographic basin because of convergence problem
when the interpolation was done at the France scale. We had to make a trade-off
between the number of data (to keep enough high to perform the interpolation)
and the level of dataset splitting. We then merge some hydrographic basin to end
up with four region of interpolation: West, East, South and North.

Then we modeled each variable as dependant only on the covariation between sites
and taking in account stream network. To model the autocovariance between sites,
we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their prediction performance. The interpolations
were done with the `SSN` R package 
[@isaak_applications_2014; @hoef_ssn:_2014].

@@

The performance of the interpolation was evaluated with leave one out
cross-validation (already implemented in the SSN package), i.e. the value of the
interpolated variable of each station was predicted after removing it from the
dataset.

@@

## Temporal series classification

The CV of biomass over time can cover multiple dynamics, including increases,
stable, decreases and a mix of them. Theoretically, however, the CV computed to
quantify stability is often represented with the assumption that the overall
dynamic of the biomass varies around an equilibrium. In order to approach this
condition, we kept only the stations that have met this criteria. To qualify the
biomass trajectories over time, we used the symbolic aggregated approximation
method (SAX) which is commonly used in time series analysis and implemented in
the `jmotif` R package. This method classifies time series rather than
quantifies trends and so is generic enough to encompass the diversity of time
series shapes. The classification necessitates three steps. First, the time
series are normalized (i.e. substituting the mean and divide by the standard
deviation). The second step is to reduce the dimensionality of the time series
by Piecewise Aggregation Approximation (PAA) method. Each time series is divided
in $N$ frames of equal size that allows to the reduce the dimensionality of the
time series. For each frame, we simply kept the mean value of the time series.
The last step is to qualify each frame into string. Because values of
normalized time series typically follows a normal distribution ($N(0,1)$), a
breakpoints can be computed by dividing the area under the normal distribution
in equal parts. The algorithm attributes a alphabetic letter to each part from
the lowest to the highest ($a$, $b$, $c$, ...). We set three breakpoints, so
$a$, $b$ and $c$. Finally, the trajectory of each time series is classified as
a combination of $a$, $b$, $c$ letters. $abc$ and $cba$ time series represent
respectively monotonic increasing and decreasing trajectories. For data analysis,
we kept only the $bbb$ which represent time series that showed no temporal
tendencies, with respect to our methodology.

## PCA over environmental variables

@@

```{r, cache = FALSE}
prop_var_expl_pca <- pca_rotated$rotated$Vaccounted["Proportion Var", ] %>%
  map_dbl(., ~round(.x * 100))

@@

pca_caption <- paste0(
  "Rotated axis of PCA on environmental variables. The five components
  explained,", tot_var_expl_pca, "% of the variance (",
  paste0(names(prop_var_expl_pca), ":", prop_var_expl_pca, "%"), ").",
  "(A) Axis 1 was positively related to the stream size, axis 2 was related to
  the stream slope and altitude and negatively related to average water temperature and
  BOD. (B) Eigenvalues distribution, we kept the first two axis for the data analysis. ",
  "Please note that we took the inverse of the second axis in the data analysis,
  to make the axis more easily interpretable, i.e. positively related de average
  water temperature."

  )
```

```{r pca-rotated, cache = FALSE, fig.cap = pca_caption}
p1 <- my_pca_plot(xaxis = "RC1", yaxis = "RC2", ctb_thld = .4)
#p2 <- my_pca_plot(xaxis = "RC3", yaxis = "RC4", ctb_thld = .4)
#p3 <- my_pca_plot(xaxis = "RC4", yaxis = "RC5", ctb_thld = .4)

@@

@@

```{r}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 2)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 2), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c("Log base 10", "Log base 10 and inversion", "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  interpretation = c("linked to stream size",
    "linked to average temperature and BOD",
    "Total species richness over the temporal series",
    "Median of annual link density", "Median of annual average trophic level", "", "",
    "Median of annual total biomass")) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "Variable description and transformation"
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
sem_variable_table
```

---
#### SEM hypothesis
#
#We linked environmental variables both to community structure and stability
#components. Rooted in assembly theory, environment filters species that can
#establish in a community and thus affect community structure. (+ add specific
#hypothesis about physical and climate). (1) We hypothesized that enrichment
#(characterized by DBO) and mean annual temperature will increase species turnover
#[@friedpetersen_drivers_2020]. (2) We hypothesized environmental forcing can also
#decrease biomass stability directly without going by community structure. First,
#instability of habitat and environmental factors may in turn drive biomass
#instability [@hansen_climate_2013]. Secondly, average environmental conditions
#such as enrichment can increase directly CV by increasing species biomass
#fluctuations [@tabi_warming_2019]. Increasing temperature was also hypothesized
#to decrease stability through starvation resulting from high metabolic activity
#[@tabi_warming_2019].
#
#---
## Species richness to community attributes 
#---
#
#To reveal the mechanisms behind the relationship between species richness and
#biomass stability, causal links were set from species richness to other
#community attributes. In particular, species richness decreased connectance
#following results from @thebault_stability_2010; @winemiller_must_1989. But
#theoretically, some ecological networks can show the inverse relationship
#[@winemiller_must_1989]. (6) We hypothesized also that species richness will
#modulate average weighted trophic level because of food web assembly processes.
#A community with more fish species can support higher trophic level. Finally, we
#set up causal relationships from community attributes to stability components. 
#
#---
##  Community attributes to stability
#---
#We hypothesized that (7) species richness decreased synchrony between species
#[@loreau_species_2008] and increased $\overline{CV_{sp}}$
#[@tilman_biodiversity_2006]. (8a) We expected that average trophic level had a
#positive or negative relationship with synchrony, bottom-up effects should
#result in positive whereas top-down in negative relationships. We expected a (8b)
#negative relationship between average trophic level and $\overline{CV_{sp}}$ if
#predators stabilised prey fluctuations or if the increasing part of
#biomass in the predators decreased $\overline{CV_{sp}}$ because of their higher
#body mass (if mean increase faster than variance). (9a) Connectance was expected
#to decrease or increase $\overline{CV_{sp}}$ and (9b) synchrony because of top
#down effects or bottom up effects.
---

@@

```{r sensitivity, cache = TRUE, message = FALSE, fig.dim = c(10, 7), fig.cap = "Sensivity analysis with different dataset"}
library(furrr)
